Realizing Temporal Graphs

### 2023-01-17 RealizingTemporalGraphs.tex
Changes 17th january

Observe that edges incident with $y^T, y^F$ that have label three connect $y^T, y^F$ to $v$. Furthermore, $v$ has distance two from $x$ in the unlabeled graph $G$. Since all edges incident with variable vertices $x$ have label two, there is no temporal path starting at $x$ and arriving at $v$ at time two, that could then continue to $y_T$ or $y_F$ using an edge with label three. Hence, there is no temporal path from $x$ to $y^T, y^F$ using an edge with label three. Since all edges incident with $x$ have label two and variable vertices $x$ are non-adjacent to variable vertices $y^T,y^F$, there cannot be a temporal path from variable vertex $x$ to variable vertex $y^T$ or $y^F$.
\item For all variable vertices $x$ and clause vertices $c$ we have $d(v,x)=d(v,c)=\infty$:
\todo[inline]{New matrix entries D for period Delta: d(v,x)=Delta (using two edges, with labels L+1,L). 
New matrix entries D for period Delta: d(v,c)=Delta (using two edges, with labels L,L-1).}

@@

\section{FPT and W[1]-hardness of exact realization}
In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subset E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$.
The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$.
Note that one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$, \ie $F = E(G) \setminus E(T)$.

@@

Before we start with the proof we fix the following notation. 
Let $D$ be the input matrix of \deltaExact\ \ie
the matrix of the fastest temporal paths among $n$ vertices, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
Let $F$ be a minimum feedback edge set of $G$ and let $k$ be the feedback edge number of $G$.

We now introduce an algorithm that determines the labeling of $G$, satisfying $D$, that runs in $f(k) \cdot \text{poly}(n)$ time.
Our algorithm performs the following steps.
\begin{enumerate}
    \item Preprocessing on graph $G$. Determine the set $E$ of feedback edges and the set $U$ of vertices of interest, where $|E|, |U| \in O(k)$.
    \item Guess the fastest temporal paths among vertices in $U$, for each guess create an ILP with variables corresponding to fastest temporal paths.
    All in total create $f(k)$ partial ILP instances, each with $f(k)$ variables
    \item For each partial ILP instance I:
    \begin{enumerate}
        \item create $f(k)$ final instances $I_I$,
        \item check if $I_I$ has a solution.
    \end{enumerate}
\end{enumerate}
\begin{comment}
\begin{enumerate}
    \item Create $G'$ and $D'$ from $G$ and $D$ respectively, by removing all degree $1$ vertices from $G$ and $D$.
    \item Compute a minimum feedback edge set $F$ in $G'$ (note, $F$ is also a minimum feedback edge set of $G$).
    \item Determine vertices of interest $U$ of the graph $G'$.
    \item \label{enum:FPT-labelingVerticesOfInterest} Guess the (partial) labeling $\lambda$ of $G'$, that connects all pairs of vertices from $U$, satisfying their respective values in $D'$.
    \item \label{enum:FPT-labelingraphNoLeafs} Extend the partial labeling $\lambda$ to the labeling of the whole $G'$, respecting $D'$.
    \item \label{enum:FPT-labelingWholeGraph} Extend the labeling $\lambda$ of $G'$ to the labeling of $G$, respecting $D$.
\end{enumerate}

It is important to point out, that in step~\ref{enum:FPT-labelingVerticesOfInterest}, our algorithm does not build a labeling satisfying conditions in this step,
and then tries to extend it in steps~\ref{enum:FPT-labelingraphNoLeafs} and \ref{enum:FPT-labelingWholeGraph},
but it
builds (many different instances of) integer linear program (ILP) problem with $O(k)$ variables,
with restrictions satisfying condition from steps~\ref{enum:FPT-labelingVerticesOfInterest}--\ref{enum:FPT-labelingWholeGraph}.
At the end we solve the ILP problem on all instances and if there is a positive solution, it gives rise to the desired labeling $\lambda$ of $G$.
For each of the steps we now argue its correctness and time complexity.
\end{comment}

\subsection{Step 1 of the FPT algorithm}
In the first step of the algorithm we create an instance $D'$, with the underlying graph $G'$, from $D$ and $G$, by iteratively removing vertices of degree one from $G$ and their corresponding rows and columns from the matrix $D$. 
This part is executed in $O(n+m)$ time.

Then we determine a minimum feedback edge set $F$ of $G'$ by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
Lastly, we determine the set $U$ of \emph{vertices of interest}.
Let $T$ be a spanning tree of $G'$, with $F$ being the corresponding feedback edge set of $G'$.
Let $V_1 \subset V(G')$ be the set of leaves in $T$, $V_2 \subset V(G')$ be the set of vertices of degree two in $T$, that are incident to some edges in $F$, and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ are incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with $U = V_1 \cup V_2 \cup V_3$ the set of vertices of interest. It follows that $|U| \leq 4k$.
We determine $F$ and $U$ in linear time.

\subsection{Step 2 of the FPT algorithm}
In this step we want to guess the structure of fastest paths (underlying paths) among all pairs of vertices of $U$. %and then determine their labeling.
For a pair $u,v$ of vertices in $U$, there are $k!$ possible paths in $G'$ between them. 
Therefore, we have to try all $O(k^k)$ possible paths, where at least one of them will be a fastest temporal path from $u$ to $v$.
Repeating this for all pairs of vertices $u,v \in U$ we get $O((k^k)^{k^2})=O(k^{k^3})$ different variations of fastest temporal paths among all vertices.

Each partial ILP instance is connected to one list of specific fastest temporal paths among the vertices from $U$.
From now on, for every par of vertices $u,v \in U$ we assume that we know exactly what is the underlying path of a fastest temporal path from $u$ to $v$.
We describe now how to build a specific partial ILP instance from a specific list of fastest temporal paths.
While building the ILP instances we exploit the following property greatly, that is why we state it as a separate observation.

@@

In the following we prove that knowing the structure (the exact underlying path) of a fastest temporal path $P$, among two vertices of interest $u,v \in U$,
results in determining the labeling of each edge in the fastest temporal path from $u$ to $v$ (with the exception of some constant number of edges), with respect to the label of the first edge.

\begin{claim}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1,v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Suppose also that no other vertex from $P$ is a vertex of interest, \ie $P \setminus \{u,v\} \cap U = \emptyset$.
    We can determine the labeling $\lambda$ of every edge in $P$, up to the constant $c_{uv} = \lambda(uv_2)$,
    such that the labeling $\lambda$ respects $D'$.
\end{claim}

@@

\begin{claim}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1, v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Let $\ell_{u,v} \geq 1$ be the number of vertices of interest in $P$ different to $u,v$, namely $\ell_{u,v} = | v_i \in \{P \setminus \{u,v \} \} \cap U |$.
    We can determine the labeling $\lambda$ of all but $\ell_{u,v}$ edges of $P$, up to the constant $c_{uv} = \lambda(uv_2)$,
    such that the labeling $\lambda$ respects $D'$.
    Moreover, for the labels of edges that cannot be determined precisely, we can get some extra restrictions.
\end{claim}

\begin{proof}
We denote with $v_i \in U$ a vertex of interest in $P \setminus \{u,v\}$.
There are two options, either $v_i$ is reached from $u$ using the subpath $P_i = (u, v_2, v_3, \dots, v_i)$ of $P$,
or there exists a fastest temporal path $P' = (u, w_2, w_3, \dots, w_{p'} = v_i)$, for which $P \cap P' = \{u, v_i\}$.
If the temporal path from $u$ to $v_i$ is a subpath of $P$, then we determine the labeling the same as in previous case.
So we can suppose that the fastest temporal path from $u$ to $v_i$ is of form $P'$.\todo{TBD - George prooved it}
\end{proof}

We use the above results in the following way.
For all vertices of interest $u \in U$ and $v \in U \setminus \{v\}$, we know the underlying path $P_{u,v}$ of a fastest temporal path from $u$ to $v$. 
We create a variable $c_{u,v}^{e_1} \in [\Delta]$ that corresponds to the label of the first edge of $P_{u,v}$ in $(G', \lambda')$ and potentially at most $\ell_{u,v}$ variables $c_{u,v}^{e_w} \in [\Delta]$ corresponding to 
labels of edges $e_w \in P_{u,v}$ that are incident to a vertex of interest $w$, for which the underlying path $P_{u,w}$ of a fastest temporal path from $u$ to $w$ intersects $P_{u,v}$ only in vertices $u$ and $w$.
Labels on all other edges of $P_{u,v}$ are determined exactly, in relation to the constant $c_{u,v}^{e_1}$.
Repeating this procedure for all $u,v \in U$ creates at most $|U|^3 = O(k^3)$ new variables.

\todo[inline]{Write example for variables and constrains this part produces.}

\subsection{Step 3 of the FPT algorithm}

After finishing the second step of the FPT algorithm labels on all edges of fastest temporal paths from $u$ to $v$, where $u,v \in U$, are determined with respect to the label of the first edge.
There may still be some edges in the graph, that are unlabeled. Note that these edges are exactly the ones that are not a part of any fastest temporal path.
Let us first observe some properties of such edges.

\todo[inline]{Fix the following proof, add more details/definitions to the introduction.}
\begin{claim}
    Unlabeled edges form disjoint paths in $G'$, where each path $P$ starts and ends in some vertex of interest, 
    \ie for every unlabeled path $P$ there exist $u,v \in U$, such that $P = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}

\begin{proof}
    We prove this claim in two steps.
    First, all inner vertices of $P$, \ie vertices $v_i \in P \setminus \{u,v\}$ are of degree exactly $2$. They are of degree at least $2$ since they are on a path $P$ and are incident to two unlabeled edges in  $P$.
    They are of degree at most $2$ as if they were of a higher degree they would be vertices of interest and in this case we could split $P$ into shorter (sub)paths.

    Second, suppose that $P$ does not start in a vertex of interest, therefore $u \notin U$. As $u$ is the starting vertex of the unlabeled path $P$, there must exist a vertex of interest $u'$ such that a path between $u$ and $u'$ is labeled.
    Since $u$ is not a vertex of interest, the only reason to label edges between $u'$ and $u$ would be if they were a part of some fastest temporal path among some vertices of interest and $u$ would be an inner vertex on that path.
    Therefore $u$ would have to be incident to two labeled edges (that are part of some fastest temporal path) and one unlabeled edge, but in this case the degree of $u$ would be at least $3$ and therefore $u$ would be a vertex of interest.
\end{proof}

\begin{claim}
    Among any pair of vertices of interest $u,v \in U$ there can be at most one unlabeled path $P = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}
\begin{proof}
    Suppose for the contradiction that there are two unlabeled paths among $u,v \in U$.
    Denote them with $P$ and $P'$. 
    Then $C = P \cup P'$ form a cycle in $G'$. Since $G'$ admits no loops, the cycle $C$ is of length at least $3$.
   Since $C$ is a cycle, at least one of it's edges $e \in E(C)$ has to be in the feedback edge set $F$, and therefore, by the definition of vertices of interest $U$, 
   both endpoints of $e$ have to be in $U$.
   So $P$ and $P'$ cannot be both fully unlabeled.
\end{proof}

From the above it follows that there are at most $f(k)$ (namely $O(|U|^2$) unlabeled paths in $G'$ after algorithm finishes Step 2.
Now we want to extend the labeling from Step 2 to all unlabeled paths.
To do this we want to relate the labeling of an unlabeled path to some labeled (\ie fastest temporal) path.
Let us first observe some properties.

Suppose that $Q = (u,v_2,v_3, \dots, v_p, v)$, where $p \geq 2$ is an unlabeled path in $G'$, between vertices of interest $u,v \in U$,
let $P$ be the underlying path of a fastest temporal path from $u',v' \in U$.
Let $P$ be such path that there exists a vertex $x \in P \setminus \{u,v\}$ of degree two (\ie $x\notin U$).

@@

    Suppose that $Q = (u,v_2,v_3, \dots, v_p, v)$, where $p \geq 2$ is an unlabeled path in $G'$, between vertices of interest $u,v \in U$,

@@

To calculate $\lambda (s,v_i)$ we need $O(|N_{G'}(u)|) = O(n)$ time, therefore to calculate all labels of edges incident to vertices of degree one in $G$, we need $O(n^2)$ time, calculating $D'$ takes $O(n^3\log n)$ time (see \cref{lemma:calculatingD}), therefore the whole step is performed in polynomial time.
\end{comment}
%
%
\subsubsection*{Step 3(a) of the FPT algorithm}
Using all of the above results we proceed in the following way.
For a partial ILP instance $I$ we 
select an inner vertex $x$ of every fastest temporal path among vertices of interest $u'$ and $v'$,
and for every unlabeled path among vertices of interest $u$ and $v$ we find such $y$ and $y_1, y_2$ from this path, that $y$ is furthest away from $x$ and $y_1, y_2$ are its neighbors.
We then enumerate $4$ possible ways that a fastest temporal path starts from $x$ and reaches $y$, 
and then $2$ additional ways, how a fastest temporal path starts from $x$ to $y_1$ (resp.~$y_2$).
For each of the possibilities we then precisely determine labels of the unlabeled path, up to the constant of the fastest $u',v'$ temporal path.

Since there are at most $|U|^2$ fastest temporal paths among vertices of interest, we create at most $4|U|^2$ extra instances for each unlabeled path.
Therefore one partial ILP instance $I$ is further extended into $O(|U|^3)=f(k)$ \todo{$O(2^k)$} final ILP instances $I_I$.

\subsubsection*{Step 3(b) of the FPT algorithm}
Using results from \cite{Lenstra1983Integer} we know that an ILP instance can be solved in FPT time with respect to the number of variables. In the case when the input instance as a solution, algorithm also provids us with one.

>>>>

Observe that edges incident with $y^T, y^F$ that have label three connect $y^T, y^F$ to $v$. Furthermore, $v$ has distance two from $x$ in the undetermined graph $G$. Since all edges incident with variable vertices $x$ have label two, there is no temporal path starting at $x$ and arriving at $v$ at time two, that could then continue to $y_T$ or $y_F$ using an edge with label three. Hence, there is no temporal path from $x$ to $y^T, y^F$ using an edge with label three. Since all edges incident with $x$ have label two and variable vertices $x$ are non-adjacent to variable vertices $y^T,y^F$, there cannot be a temporal path from variable vertex $x$ to variable vertex $y^T$ or $y^F$.
\item For all variable vertices $x$ and clause vertices $c$ we have $d(v,x)=d(v,c)=\infty$:
\todo[inline]{New matrix entries D for period Delta: d(v,x)=Delta (using two edges, with labels L+1,L). 
New matrix entries D for period Delta: d(v,c)=Delta (using two edges, with labels L,L-1).}

@@

\section{FPT and W[1]-hardness of exact realization -- new version}
In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subset E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$.
The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$.
Note that one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$, \ie $F = E(G) \setminus E(T)$.

@@

Before we start with the proof of \cref{thm:FPTwrtFES} we fix the following notation. 
Let $D$ be the input matrix of \deltaExact\ \ie
the matrix of the fastest temporal paths among $n$ vertices, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
Let $F$ be a minimum feedback edge set of $G$ and let $k$ be the feedback edge number of $G$.

Throughout the proof, whenever we say \emph{guess} the path/subpath/vertex, etc., we mean that we enumerate all possible solutions.
For example, let $u$ and $v$ be two vertices in a graph $G$. 
We say that we \emph{guess} that $P$ is the underlying path of a fastest temporal path $(P, \lambda)$ from $u$ to $v$ in $(G, \lambda)$.
This means that we iterate over all possible possible $u,v$-paths $P_1, P_2, \dots $ in $G$, 
and form multiple cases, 
one in which $(P_1, \lambda)$ is a fastest temporal path from $u$ to $v$ in $(G,\lambda)$,
one when $(P_2, \lambda)$ is a fastest temporal path from $u$ to $v$ in $(G,\lambda)$, etc.
To correctly determine if the problem has a valid solution, we must check all cases.
In each case we perform required calculations, 
and we either come to the contradiction (meaning that the current guess was wrong) 
or our assumption holds (meaning that the guessed path is really a fastest path from $u$ to $v$ in $(G, \lambda)$).

Note that we can easily solve the \deltaExact\ by guessing the structure of temporal paths among all pairs of vertices.
But, this would create $f(n,m)$ guesses, and the running time of the algorithm would be too big.
Therefore, we approach the problem using some special properties, which allow us to create $f(k)$ different guesses.
We solve each guess by creating an ILP instance with $f(k)$ variables and $O(n^c)$ constraints, for some $c \in \mathbb{N}$.
The ILP instance is a \textsc{Yes} instance if and only if $D$ is a \textsc{Yes} instance of \deltaExact.
Moreover, from a positive solution of the ILP instance we can construct the desired labeling $\lambda$.
Using results from \cite{Lenstra1983Integer} we know that an ILP instance can be solved in FPT time with respect to the number of variables.
Therefore we can solve our ILP instance in $f(k)$ time.

We now introduce the algorithm, that determines the labeling of $G$, satisfying $D$, and runs in $f(k) \cdot \text{poly}(n)$ time.
Our algorithm performs the following steps.
\begin{enumerate}
    \item Preprocessing on graph $G$. Determine the set $E$ of feedback edges, the set $U$ of vertices of interest and the set $U^*$ of neighbors of vertices of interest, where $|E|, |U|, |U^*| \in O(k)$.
    \item Guess what are the fastest temporal paths among vertices in $U$ and 
    what are the fastest temporal paths among the vertices in $U^*$, that are of some specific form.
    \begin{enumerate}
        \item For each guess create an ILP instance $I$ with $f(k)$ variables.
        \item Add constraints for all pairs of vertices.
    \end{enumerate}
    \item Solve the ILP instance $I$ and check if a solution exists. In the positive case create its corresponding labeling $\lambda_I$ and check if it satisfies $D$.
    All in total we create $f(k)$ partial ILP instances, each with $f(k)$ variables and $f(n)$ constraints.
\end{enumerate}

\subsection{Step 1 of the FPT algorithm}
From the underlying graph $G$ of $D$ we first create a graph $G'$ by
iteratively removing vertices of degree one from $G$.
Then we determine a minimum feedback edge set $F$ of $G'$ by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
Note that $F$ is also a minimum feedback edge set of the whole $G$.
Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, with $F$ being the corresponding feedback edge set of $G'$.
Let $V_1 \subset V(G')$ be the set of leaves in the spanning tree $T$, $V_2 \subset V(G')$ be the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with $U = V_1 \cup V_2 \cup V_3$ the set of vertices of interest. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie $U^* = \{v \in V(G') \setminus U | u \in U, v \in N(u)\}$.
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, therefore $U^* \in O(k^2)$.

The whole Step 1 is performed in $O(n + m)$ time.

\subsection{Step 2 of the FPT algorithm}
Let us denote with $Z$ the set of vertices removed from $G$ at the beginning of the algorithm, \ie $Z = V(G) \setminus V(G')$.
By the construction it follows that $Z$ consists of disjoint trees. 
Note that the fastest temporal paths among vertices of $G'$ are independent of vertices in $Z$.
Let us remember the following basic property of the problem \deltaExact.
A labeling $\lambda$ of $G$ satisfies $D$ if every fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$,
and none other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases,
where the first vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the last vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases we guess the underlying path $P$ of the fastest temporal path from the vertex $v_i$ to $v_j$, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(n,m)$ constraints. We have to guarantee only, that the number of guesses, of faster temporal paths, is in $f(k)$ and not in $f(n)$.

We start the algorithm by guessing the structure of 
the fastest temporal paths (we guess the underlying paths) among all pairs of vertices of $U$. 
For a pair $u,v$ of vertices in $U$, there are $k!$ possible paths in $G'$ between them. 
Therefore, we have to try all $O(k^k)$ possible paths, where at least one of them will be a fastest temporal path from $u$ to $v$, respecting the values from $D$.
If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
Repeating this procedure for all pairs of vertices $u,v \in U$ we get $O((k^k)^{k^2})=O(k^{k^3})$ different variations of the fastest temporal paths among all vertices in $U$.

We now create one partial ILP instance for each guess of fastest temporal paths among the vertices from $U$,
which we later on extend to $f(k)$ final ILP instances, after guessing some extra fastest temporal paths.
From now on, for every pair of vertices $u,v \in U$ we assume that we know exactly what is the underlying path of a fastest temporal path from $u$ to $v$.

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$ in $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment}. Denote it with $S_{u,v}$.
\end{definition}

We describe now how to build one partial ILP instance from a specific list of fastest temporal paths.
While building the ILP instances we exploit the following property greatly, that is why we state it as a separate observation.

@@

In the following we prove that knowing the structure (the underlying path) of a fastest temporal path $P$, from a vertex of interest $u$ to a vertex of interest $v$,
results in determining the labeling of each edge in the fastest temporal path from $u$ to $v$ 
(with the exception of some constant number of edges), with respect to the label of the first edge.
More precisely, if path $P$ from $u$ to $v$ is a segment, then we can determine labels of all edges,
and if $P$ consists of $\ell$ segments, then we can determine the labels of all but $\ell -1$ edges.
For the exact formulation and proofs see \cref{claim:FPT-uv-Labelalledges,claim:FPT-uv-LabelAlmostalledges}.

\begin{claim}\label{claim:FPT-uv-Labelalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1,v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Suppose also that no other vertex from $P$ is a vertex of interest, \ie $P \setminus \{u,v\} \cap U = \emptyset$.
    We can determine the labeling $\lambda$ of every edge in $P$, up to the constant $c_{uv} = \lambda(uv_2)$,
    such that the labeling $\lambda$ respects $D'$.
\end{claim}

@@

\begin{claim}\label{claim:FPT-uv-LabelAlmostalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1, v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Let $\ell_{u,v} \geq 1$ be the number of vertices of interest in $P$ different to $u,v$, namely $\ell_{u,v} = | v_i \in \{P \setminus \{u,v \} \} \cap U |$.
    We can determine the labeling $\lambda$ of all but $\ell_{u,v}$ edges of $P$, up to the constant $c_{uv} = \lambda(uv_2)$,
    such that the labeling $\lambda$ respects $D'$.
    Moreover, for the labels of edges that cannot be determined precisely, we can get some extra restrictions.
\end{claim}

\begin{proof}
We denote with $v_i \in U$ a vertex of interest in $P \setminus \{u,v\}$.
There are two options, either $v_i$ is reached from $u$ using the subpath $P_i = (u, v_2, v_3, \dots, v_i)$ of $P$,
or there exists a fastest temporal path $P' = (u, w_2, w_3, \dots, w_{p'} = v_i)$, for which $P \cap P' = \{u, v_i\}$.
If the temporal path from $u$ to $v_i$ is a subpath of $P$, then we determine the labeling of $P$ using \cref{claim:FPT-uv-Labelalledges}.
So we can suppose that the fastest temporal path from $u$ to $v_i$ is of form $P'$.\todo{TBD - George proved it}
\end{proof}

We use the above results in the following way.
Let $u, v \in U$ be two vertices of interest. 
We know the underlying path $P_{u,v}$ of a fastest temporal path from $u$ to $v$. 
Suppose that $P_{u,v}$ consist of $\ell$ segments $S_{u_i,v_i}$.
Using \cref{claim:FPT-uv-LabelAlmostalledges} we determine the labels of all but $\ell -1$ edges on $P_{u,v}$, with respect to the value of the first edge.
This equals to creating a constraint by setting the duration of the temporal path along $P_{u,v}$ to equal the value $D_{u,v}$.
There are $f(k)$ different paths $P^i_{u,v}$ from $u$ to $v$ in $G$.
We now have to iterate through all of them and add the constraint that the duration of $(P^i_{u,v}, \lambda)$ is greater or equal to $D_{u,v}$.
We repeat the above procedure for all pairs of vertices in $U$.
This results in producing $f(k)$ variables and constraints for our partial ILP.

\begin{comment}
We create a variable $c_{u,v}^{e_1} \in [\Delta]$ that corresponds to the label of the first edge of $P_{u,v}$ in $(G', \lambda')$ and potentially at most $\ell_{u,v}$ variables $c_{u,v}^{e_w} \in [\Delta]$ corresponding to 
labels of edges $e_w \in P_{u,v}$ that are incident to a vertex of interest $w$, for which the underlying path $P_{u,w}$ of a fastest temporal path from $u$ to $w$ intersects $P_{u,v}$ only in vertices $u$ and $w$.
Labels on all other edges of $P_{u,v}$ are determined exactly, in relation to the constant $c_{u,v}^{e_1}$.
Repeating this procedure for all $u,v \in U$ creates at most $|U|^3 = O(k^3)$ new variables.
\end{comment}
\todo[inline]{Write examples for variables and constrains this part produces.}

After finishing the previous step for all pairs of vertices in $U$,
the edges of fastest temporal paths from $u$ to $v$, where $u,v \in U$, are determined with respect to the label of the first edge.
Let us call these edges \emph{determined edges}, and the edges that were not determined with the respect to some other edge label up to now the \emph{undetermined} edges.
Note that these edges are exactly the ones that are not a part of any fastest temporal path.
Let us first observe some properties of such edges.

\todo[inline]{Fix the following proof, add more details/definitions to the introduction.}
\begin{claim}
    Undetermined edges form disjoint segments in $G'$, where each segment starts and ends in some vertex of interest, 
    \ie for every undetermined path $P$ there exist $u,v \in U$, such that $P = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}
\begin{proof}
    We prove this claim in two steps.
    First, all inner vertices of $P$, \ie vertices $v_i \in  S \setminus \{u,v\}$ are of degree exactly $2$. They are of degree at least $2$ since they are on a path $S$ and are incident to two undetermined edges in  $P$.
    They are of degree at most $2$ as if they were of a higher degree they would be vertices of interest and in this case we could split $P$ into shorter segments (subpaths).

    Second, suppose that $P$ does not start in a vertex of interest, therefore $u \notin U$. As $u$ is the starting vertex of the undetermined path $P$, there must exist a vertex of interest $u'$ such that a path between $u$ and $u'$ is labeled.
    Since $u$ is not a vertex of interest, the only reason to label edges between $u'$ and $u$ would be if they were a part of some fastest temporal path among some vertices of interest and $u$ would be an inner vertex on that path.
    Therefore $u$ would have to be incident to two labeled edges (that are part of some fastest temporal path) and one undetermined edge, but in this case the degree of $u$ would be at least $3$ and therefore $u$ would be a vertex of interest.

    It follows that $P$ is a segment.
\end{proof}

\begin{claim}
    Among any pair of vertices of interest $u,v \in U$ there can be at most one undetermined segment $S = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}
\begin{proof}
   Suppose for the contradiction that there are two undetermined segments among $u,v \in U$.
   Denote them with $S_{u,v}$ and $S'_{u,v}$. 
   Then $C = S_{u,v} \cup S_{u,v}'$ form a cycle in $G'$. Since $G'$ admits no loops, the cycle $C$ is of length at least $3$.
   Since $C$ is a cycle, at least one of its edges $e \in E(C)$ has to be in the feedback edge set $F$, and therefore, by the definition of vertices of interest $U$, 
   both endpoints of $e$ have to be in $U$.
   So $S_{u,v}$ and $S_{u,v}'$ cannot be both fully undetermined.
\end{proof}

From the above it follows that there are at most $f(k)$ (namely $O(|U|^2$) undetermined segments in $G'$ after algorithm finishes going through fastest temporal paths among vertices of interest.

We now continue with guessing fastest temporal paths among all vertices in $U^*$, 
which by similar arguing as for vertices in $U$, gives us $O(k^{k^6})$ guesses.
Besides that we also guess fastest temporal paths of the following form.
Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be two segments,
first one between vertices $u,v \in U$, second one between vertices $w, z \in U$, 
where the first one includes $v_2, v_{p-1} \in U^*$, 
and the second one includes $z_2, z_{p-1} \in U^*$.
We want to guess the underlying paths of the fastest temporal paths from $v_2$ to $z_2$,
where the first path goes 
from $v_2$, through $u$, through $w$ and then to $z_2$,
and the second one goes
from $v_2$, through $u$, through $z$ and then to $z_2$.
And two similar paths from $v_2$ to $z_{p-1}$.
We repeat this procedure for all pairs of segments and all vertices from $U^*$ in those segments.
There are $f(k)$ possible paths of this form, since there are $f(k)$ segments in $G'$.
All together we made $f(k)$ new guesses for the structure of the fastest temporal paths.
Each of the previous partial ILP instances is therefore extended into $f(k)$ new ILP instances.

We now do the following.
Let $S^i_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ be a line segment in $G'$, let us
fix a vertex $v_i \in S^i_{u,v} \setminus \{u,v\}$.
Let 
$S^j_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be another segments in $G'$.
Now we check the matrix $D$ and determine which vertex $z_j \in S^j_{w,z} \setminus \{w,z\}$ 
has the biggest value $D_{v_i,z_j}$.
Let us call this vertex a \emph{split vertex of $v_i$ in the segment $S_{wz}$}.
Denote with $z_{j-1}$ and $z_{j-1}$

Let us first observe some properties.

Suppose that $Q = (u,v_2,v_3, \dots, v_p, v)$, where $p \geq 2$ is an undetermined segment in $G'$, between vertices of interest $u,v \in U$,
let $P$ be the underlying path of a fastest temporal path from $u',v' \in U$.
Let $P$ be such path that there exists a vertex $x \in P \setminus \{u,v\}$ of degree two (\ie $x\notin U$).

@@

    Suppose that $Q = (u,v_2,v_3, \dots, v_p, v)$, where $p \geq 2$ is an undetermined segment in $G'$, between vertices of interest $u,v \in U$,

@@

To calculate $\lambda (s,v_i)$ we need $O(|N_{G'}(u)|) = O(n)$ time, therefore to calculate all labels of edges incident to vertices of degree one in $G$, we need $O(n^2)$ time, calculating $D'$ takes $O(n^3\log n)$ time (see \cref{lemma:calculatingD}), therefore the whole step is performed in polynomial time.
\end{comment}
%
%
\subsubsection*{Step 3(a) of the FPT algorithm}
Using all of the above results we proceed in the following way.
For a partial ILP instance $I$ we 
select an inner vertex $x$ of every fastest temporal path among vertices of interest $u'$ and $v'$,
and for every endetermined path among vertices of interest $u$ and $v$ we find such $y$ and $y_1, y_2$ from this path, that $y$ is furthest away from $x$ and $y_1, y_2$ are its neighbors.
We then enumerate $4$ possible ways that a fastest temporal path starts from $x$ and reaches $y$, 
and then $2$ additional ways, how a fastest temporal path starts from $x$ to $y_1$ (resp.~$y_2$).
For each of the possibilities we then precisely determine labels of the undetermined path, up to the constant of the fastest $u',v'$ temporal path.

Since there are at most $|U|^2$ fastest temporal paths among vertices of interest, we create at most $4|U|^2$ extra instances for each undetermined path.
Therefore one partial ILP instance $I$ is further extended into $O(|U|^3)=f(k)$ \todo{$O(2^k)$} final ILP instances $I_I$.

\subsubsection*{Step 3(b) of the FPT algorithm}
Using results from \cite{Lenstra1983Integer} we know that an ILP instance can be solved in FPT time with respect to the number of variables. In the case when the input instance has a solution, algorithm also provides us with one.
